---
title: "R for Data Science"
output: html_notebook
---

This is my work following [R for Data science](https://r4ds.had.co.nz/), by Chadley Wickham and Garret Grolemund.

# 1 - Introduction

Install required packages:

```{r}
install.packages(c("tidyverse","nycflights13", "gapminder", "Lahman"))
```

Useful for recreating data:

```{r}
dput(mtcars)
```

Well, that was easy!

# 2 - Data visualization

```{r}
library(tidyverse)
```

## First steps

```{r}
mpg
```

Useful basic functions:

```{r}
length(mpg)

nrow(mpg)
ncol(mpg)
glimpse(mpg)
```

## Aesthetic mappings

```{r}
ggplot(mpg) + 
  geom_point(aes(x=displ, y=hwy, color=year), position="jitter" )
```

## Facets

```{r}
ggplot(mpg) + 
  geom_point(aes(cty, hwy), position="jitter") +
  facet_wrap(~ year)
```

```{r}
ggplot(mpg) + 
  geom_point(aes(cty, hwy), position="jitter") +
  facet_grid(year ~ fl)
```

## Geometric objects


```{r}
ggplot(mpg) + 
  geom_smooth(aes(cty, hwy), position="jitter") +
  facet_grid(year ~ fl)
```


## Statistical transformations



```{r}
ggplot(diamonds) + 
  stat_summary(aes(cut, depth), fun=median)
```


## Position adjustments

Fill vs color

"identity", "dodge", "fill", "jitter"


## Coordinate systems

Me lo fumo un poco.

## The layered grammar of graphics

```
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
```

# Workflow: basics

`Alt + Shift + K`: bring up the keyboard shortcut reference.


