---
title: "R for Data Science"
output: html_notebook
---

This is my work following [R for Data science](https://r4ds.had.co.nz/), by Chadley Wickham and Garret Grolemund.

# 1 - Introduction

Install required packages:

```{r}
install.packages(c("tidyverse","nycflights13", "gapminder", "Lahman"))
```

Useful for recreating data:

```{r}
dput(mtcars)
```

Well, that was easy!

# 2 - Data visualization

```{r}
library(tidyverse)
```

## First steps

```{r}
mpg
```

Useful basic functions:

```{r}
length(mpg)

nrow(mpg)
ncol(mpg)
glimpse(mpg)
```

## Aesthetic mappings

```{r}
ggplot(mpg) + 
  geom_point(aes(x=displ, y=hwy, color=year), position="jitter" )
```

## Facets

```{r}
ggplot(mpg) + 
  geom_point(aes(cty, hwy), position="jitter") +
  facet_wrap(~ year)
```

```{r}
ggplot(mpg) + 
  geom_point(aes(cty, hwy), position="jitter") +
  facet_grid(year ~ fl)
```

## Geometric objects


```{r}
ggplot(mpg) + 
  geom_smooth(aes(cty, hwy), position="jitter") +
  facet_grid(year ~ fl)
```


## Statistical transformations



```{r}
ggplot(diamonds) + 
  stat_summary(aes(cut, depth), fun=median)
```


## Position adjustments

Fill vs color

"identity", "dodge", "fill", "jitter"


## Coordinate systems

Me lo fumo un poco.

## The layered grammar of graphics

```
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
```

# Workflow: basics

`Alt + Shift + K`: bring up the keyboard shortcut reference.


```{r}
seq(1, 234, length.out=7)
```

# Data transformation

```{r}
library(nycflights13)
library(tidyverse)

```

## `filter`

`filter`, `arrange`, `select`, `mutate`, `summarise`


```{r}
flights %>%
  filter(carrier == "DL") %>% 
  mutate(sched_hour = floor(sched_dep_time / 100)) %>% 
  arrange(desc(dep_delay))
```

#### Exercises

(5.2.4)

1. Find all flights that

  1. Had an arrival delay of two or more hours

```{r}
flights %>% 
  filter(arr_delay > 120)
```

  2. Flew to Houston (IAH or HOU)

```{r}
flights %>% 
  filter(dest == "IAH" | dest == "HOU")
```
  
  3. Were operated by United, American, or Delta
  
  4. Departed in summer (July, August, and September)
  
  5. Arrived more than two hours late, but didn’t leave late


```{r}
flights %>% 
  filter(arr_delay > 120 & dep_delay < 15)
```
  
  6. Were delayed by at least an hour, but made up over 30 minutes in flight
  
  7. Departed between midnight and 6am (inclusive)

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

```{r}
flights %>% 
  filter(between(dep_time, 0, 600) )
```


3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
flights %>% 
  filter(is.na(dep_time)) 
```
Cancelled flights

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)


## `arrange`


```{r}
flights %>% 
  mutate(speed = distance / air_time) %>%
  arrange(desc(speed))
```


```{r}
ggplot(flights) + 
  geom_smooth(aes(x=distance, y=air_time))
```

How would I plot only those that deviate from the trend a lot (ie plotting a different dataset on the same graph?)

```{r}
ggplot(flights) + 
  geom_smooth(aes(x=distance, y=air_time)) +
  geom_point(data=filter(flights, (distance / air_time) > 8.5), aes(x=distance, y=air_time), position="jitter")
```

### `select`


```{r}
flights %>% 
  select(-(year:arr_delay))
```



There are a number of helper functions you can use within select():

* starts_with("abc"): matches names that begin with “abc”.
* ends_with("xyz"): matches names that end with “xyz”.
* contains("ijk"): matches names that contain “ijk”.
* matches("(.)\\1"): selects variables that match a regular expression. This one matches any variables that contain repeated characters. You’ll learn more about regular expressions in strings.
* num_range("x", 1:3): matches x1, x2 and x3.

`rename()` for renaming variables.

`everything()`

### `mutate`

`transmute()` keeps only the new variables.

Useful transformations:

* `%/%`: integer division
* `%%`: modulo.
* logs
* `lead()` and `lag()`.
* `cumsum` and `cummean`
* `min_rank`, `row_number`, `dense_rank`, `percent_rank`, `cume_dist`, `ntile`


### `summarise`

`mean`, `median`, `n()`, `min()`, `quantile`, `max`, `first`, `nth`, `last`, `n_distinct`, counts and means of booleans.

Subsetting:

```{r}
flights %>% 
  filter(!is.na(dep_delay)) %>%
  group_by(year, month, day) %>%
  summarise(avg_delay=mean(dep_delay),
            avg_pos_delay=mean(dep_delay[dep_delay>0]))
```

#### Ungrouping

With `ungroup`.
