---
title: "R for Data Science"
output:
  html_document:
    df_print: paged
---

This is my work following [R for Data science](https://r4ds.had.co.nz/), by Chadley Wickham and Garret Grolemund.

# 10 - Tibbles

Install required packages:

```{r}
install.packages(c("tidyverse","nycflights13", "gapminder", "Lahman"), repos = "http://cran.us.r-project.org")
library(tidyverse)
```


```{r}
as_tibble(iris)
```

```{r}
tibble(x = 1:5, 
       y = c(2,3,1,2,3),
       z = x ^ 2 + y
       )
```

Tribbles are another variant that is designed for data entry.




```{r}
nycflights13::flights %>% 
  sample_n(10) %>%
  print(n = 10, width = Inf)
```

`options(tibble.width = Inf)`


```{r}
df = tibble(
  x = rnorm(10),
  y = runif(10)
)
df %>%
  .$y
```



```{r}
typeof(df[["mistake"]])
typeof(df[["x"]]) == typeof(df$x)
typeof(df[1])
typeof(df[[1]])
```

# 11 - Data import

## The `parse_*` functions

`problems()`

```{r}
x = parse_integer(c("1", "52", "holi"))
x
```

`guess_encoding()`

### `parse_date`, `parse_datetime`, `parse_time`

Formats and ISO-8601

## Parsing a file

`guess_parser` + `parse_guess`

When guessing doesn't work:

`col_types = cols(x = col_double(), y = col_date())`

`col_types = cols(.default = col_character())`

# Tidy data

* Each variable must have its own column.
* Each observation must have its own row.
* Each value must have its own cell.

## Pivoting

`pivot_longer()` and `pivot_wider()`.

Wider used to create new columns, longer new rows.

## `separate()` and `unite()`


## Missing values

`complete()` checks all possible combinations and returns all unique.


# 13 - Relational data

Intro to relations with the flights dataset as an example.

```{r}
library(nycflights13)
airlines
```


```{r}
airports %>%
  filter(faa == 'JFK')
```

```{r}
weather
```

```{r}
weather %>%
  count(month) %>%
    ggplot() + 
      geom_col(aes(x=month, y=n))
```

```{r}
weather %>%
  ggplot() + 
    geom_bar(aes(x=month))
```

### Exercises 13.3.1

Add a surrogate key to flights

```{r add_surrogate}
flights2 = flights %>% mutate(surrogate = paste(year, month, day, carrier, flight, origin, dest, sep='-'))

flights2 %>% 
  sample(20) %>% 
  select(surrogate)
```


## Mutating joins

Prep 

```{r}
flights2 = flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)

flights2 %>% 
  left_join(airlines)
```

Also: `inner_join()`, `left_join()`, `full_join()` 


## Join keys

* Default: `by = NULL`: all common variables
* `by = "x"`: Same but only specified columns. Adds suffixes.
* `by = c("a" = "b")`: If they have different names.



### Exercises 13.4.6

1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

```{r}
flights %>%
  group_by(dest) %>%
  summarise(percent_delayed = mean(arr_delay > 15, na.rm=TRUE),
            n = n()) %>%
  left_join(airports, by=c('dest' = 'faa')) %>%
  
  ggplot(aes(lon, lat, size=n, color=percent_delayed)) +
    borders("state") +
    geom_point() +
    coord_quickmap() + scico::scale_color_scico()
```

3. Is there a relationship between the age of a plane and its delays?

```{r}
flights %>% 
  group_by(tailnum) %>%
  summarise(av_delay = mean(dep_delay, na.rm=TRUE)) %>%
  left_join(planes) %>%
  
  ggplot() +
    geom_bin2d(aes(year, av_delay))
```

With boxplots:

```{r}
flights %>% 
  group_by(tailnum) %>%
  summarise(av_delay = mean(dep_delay, na.rm=TRUE),
            departures = n()) %>%
  filter(departures > 10) %>%
  left_join(planes) %>%
  ggplot() +
    geom_boxplot(aes(x=as_factor(year), y=av_delay))
```
Eeehhh... kinda?


What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.


```{r}
flights %>%
  filter(year==2013, month== 6, day==13) %>%
  group_by(dest) %>%
  summarise(percent_delayed = mean(arr_delay > 15, na.rm=TRUE),
            n = n()) %>%
  left_join(airports, by=c('dest' = 'faa')) %>%
  
  ggplot(aes(lon, lat, size=n, color=percent_delayed)) +
    borders("state") +
    geom_point() +
    coord_quickmap() + scico::scale_color_scico()
```

Hurricane? Google says Colorado fires, but that's not apparent from the data...


### Other implementations

`base::merge`: it's worse and slower.

Implied from the text is the fact that R can't do joins not based on equality (non-equijoins).

## Filtering joins

`semi_join()`

`anti_join()`

## Set operations

`intersect(x, y)`

`union(x, y)`: doesn't repeat rows in the intersection.

`setdiff(x, y)`

# 14 - Strings

```{r}
c(str_length('dani'), str_c('a', 'b'), str_sub('asdffsd', 2, 5))
```



## regex

```{r}
samples = c('holi', 'queriod', 'comeme', 'el', 'toto')

str_view(samples, '[aeiou]')
```
```{r}
str_view_all(samples, '[aeiou]')
```

```{r}
str_view_all(samples, '[aeiou]{2}')
```

* `{n}`: exactly n
* `{n,}`: n or more
* `{,m}`: at most m
* `{n,m}`: between n and m


### Grouping and backreferences





```{r}
str_view(fruit, '(..)\\1', match=TRUE)
```


## Tools


Match/no match:

`str_detect`

```{r}
words[str_detect(words, '^w')]
```

n_matches: `str_count`.

Extract matches: `str_extract`.

Match components: `str_match` gives them as a matrix: one column for the full match, then one for each of the groups.

`str_replace()` and `str_replace_all()`

`str_split()`


`str_view(target, pattern) == str_view(target, regex(pattern))`. `regex` has more options.

## Other uses of regular expressions

`apropos()`: searches all objects available in the environment.

`dir()`

## stringi

More advanced package with more functionality.

str_view vs stri_view

# 15 - Factors

`factor(x, levels=levels)`

`parse_factor()`

`fct_inorder()`

```{r}
gss_cat
```

```{r}
gss_cat %>%
  count(race)
```

## Modifying factor order

`fct_reorder`

_Another type of reordering is useful when you are colouring the lines on a plot. `fct_reorder2()` reorders the factor by the y values associated with the largest x values. This makes the plot easier to read because the line colours line up with the legend._

`fct_infreq`.

```{r}
gss_cat %>% 
  ggplot() +
    geom_boxplot(aes(x=fct_reorder(relig, tvhours), y=tvhours))
```


## Modifying factor levels

`fct_recode()`: Also allows grouping if you assign same new value to several old levels.

`fct_collapse()`, `fct_lump()`.

# 16 - Dates and times

```{r}
library(lubridate)
```


## Dates, datetimes, and times.

### From strings

helper parser functions: `ymd()`, `mdy()`, etc:

```{r}
ymd("2021-06-21")
```
```{r}
ymd(20210621)
```

```{r}
ymd_hm("20210621 14:00")
```

```{r}
ymd(today(), tz='UTC')
```


## From columns in a DataFrame


`make_date()` and `make_datetime()`.


```{r}
flights %>%
  select(year, month, day, hour, minute) %>%
  mutate(departure_dt = make_datetime(year, month, day, hour, minute))
```

```{r}
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))

flights_dt
```

Plot:

```{r}
flights_dt %>%
  ggplot(aes(dep_time)) + 
  geom_freqpoly(binwidth=86400)
```

### To date from datetime and vv


```{r}
as_date(now())
as_datetime(today())
```

## Components of date-times

`year`, `month`, `yday`, `wday`, `hour`, `minute`, `second`.

```{r}
flights_dt %>%
  group_by(minute(dep_time)) %>%
  summarise(avg_delay=mean(arr_delay, na.rm=TRUE)) %>%
  ggplot() + 
    geom_line(aes("minute(dep_time)", avg_delay))
```

TODO: ask someone: why does the previous one not work??

```{r}
flights_dt %>%
  mutate(minute=minute(dep_time)) %>%
  group_by(minute) %>%
  summarise(avg_delay=mean(arr_delay, na.rm=TRUE)) %>%
  ggplot() + 
    geom_line(aes(minute, avg_delay))
```

## Rounding 

`floor_date`, `round_date`, `ceiling_date`.

```{r}
left = ymd_hm('20210623 00:32')

right = ymd_hm('20210623 17:22')

floor_date(right, 'hour') - right
```


## Setting components

By using the accessor functions

```{r}
this_date = ymd('20210523')

year(this_date) = year(this_date) - 37
this_date
```

`update` allows you to copy-modify and to set several components at once.

Also useful inside mutate:


```{r}
flights_dt %>%
  mutate(dep_hour=update(dep_time, yday = 1)) %>%
  ggplot(aes(dep_hour)) + 
    geom_freqpoly(binwidth=60)
```

## Time Spans

### Durations

```{r}
this_date - update(this_date, yday=1)

class(this_date - update(this_date, yday=1))
```

```{r}
duration = as.duration(this_date - update(this_date, wday=1))
duration
```

```{r}
dseconds(12)
```

```{r}
ddays(1)
```

> Durations always record the time span in seconds. Larger units are created by converting minutes, hours, days, weeks, and years to seconds at the standard rate (60 seconds in a minute, 60 minutes in an hour, 24 hours in day, 7 days in a week, 365 days in a year).


### Periods

```{r}
seconds(12)
```

```{r}
5 * days(1)
```

Periods take into account timezones in a way that durations do not.

### Intervals

An interval is a duration *with a starting point*.

```{r}
ymd(20200228) + days(5)
```

```{r}
ymd(20200228) - years(1) + days(5)
```

```{r}
next_year = today() + years(1)

an_interval = today() %--% next_year
an_interval
```

```{r}
an_interval %/% days(1)
```


## Time zones

> In R, the time zone is an attribute of the date-time that only controls printing. For example, these three objects represent the same instant in time:

```{r}
str_subset(OlsonNames(), 'New') 
```


`with_tz` keeps the actual instant in time and shows a copy with different timezone.

`force_tz` imposes the tzone and changes the instante that the object refers to.

```{r}
birthday = ymd_hm('19840523 02:00 Europe/Madrid')

with_tz(birthday, tzone='America/New_York')
```

```{r}
force_tz(birthday, tzone='America/New_York')
```


* `c()` Will drop timezones!
