---
title: "R for Data Science"
output:
  html_document:
    df_print: paged
---

This is my work following [R for Data science](https://r4ds.had.co.nz/), by Chadley Wickham and Garret Grolemund.

# 10 - Tibbles

Install required packages:

```{r}
install.packages(c("tidyverse","nycflights13", "gapminder", "Lahman"), repos = "http://cran.us.r-project.org")
library(tidyverse)
```


```{r}
as_tibble(iris)
```

```{r}
tibble(x = 1:5, 
       y = c(2,3,1,2,3),
       z = x ^ 2 + y
       )
```

Tribbles are another variant that is designed for data entry.




```{r}
nycflights13::flights %>% 
  sample_n(10) %>%
  print(n = 10, width = Inf)
```

`options(tibble.width = Inf)`


```{r}
df = tibble(
  x = rnorm(10),
  y = runif(10)
)
df %>%
  .$y
```



```{r}
typeof(df[["mistake"]])
typeof(df[["x"]]) == typeof(df$x)
typeof(df[1])
typeof(df[[1]])
```

# 11 - Data import

## The `parse_*` functions

`problems()`

```{r}
x = parse_integer(c("1", "52", "holi"))
x
```

`guess_encoding()`

### `parse_date`, `parse_datetime`, `parse_time`

Formats and ISO-8601

## Parsing a file

`guess_parser` + `parse_guess`

When guessing doesn't work:

`col_types = cols(x = col_double(), y = col_date())`

`col_types = cols(.default = col_character())`

# Tidy data

* Each variable must have its own column.
* Each observation must have its own row.
* Each value must have its own cell.

## Pivoting

`pivot_longer()` and `pivot_wider()`.

Wider used to create new columns, longer new rows.

## `separate()` and `unite()`


## Missing values

`complete()` checks all possible combinations and returns all unique.





